A user-extensible and adaptable parser architecture

### 2008-05-28 doc/logparser-acronyms.tex
Lots of corrections after proofreading.

\newacronym{RBL}{Real-time Black List}{description={
    Real-time Black Lists are a simple collaborative anti-spam technique
    used to reject or penalise email sent from mail servers reported to
    have sent large volumes of spam.  To use an RBL the mail server makes a
    DNS request incorporating the IP address of the currently connected
    client; if the requested hostname exists the client is on the RBL, and
    the mail server can decide what course of action to take.
}}

@@

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP, originally defined in \RFC{}~821~\cite{RFC821} and updated in
    \RFC{}~2821~\cite{RFC2821}, is the protocol which transfers mail from
    the sender to the recipient across the Internet.  It is a simple, human
    readable, plain text protocol, making it quite easy to test and debug
    problems with it.  A detailed description of \SMTP{} is beyond the
    scope of this document: introductory guides can be found
    at~\cite{smtp-intro-01, smtp-intro-02}, the definitive references
    are~\cite{RFC821, RFC2821}.
    %XXX A brief introduction to SMTP is
    %provided in \sectionref{SMTP background}.
}}

@@

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a method of matching patterns against text that
    are explained in detail in~\cite{Wikipedia-regex, perlre}.
}}

@@

\newacronym{SLCT}{Simple Logfile Clustering Tool}{description={
    SLCT~\cite{slct-paper} is a tool implementing an algorithm designed by
    Risto Vaarandi~\cite{risto-vaarandi} for identifying and grouping
    similar log files, then producing a regex for each group of lines.
}}

@@

\newacronym{NLP}{Natural Language Processing}{description={
    Natural Language Processing~\cite{Wikipedia-nlp} attempts to increase
    our understanding of the languages normally used by humans (e.g.\
    English, Spanish, Japanese), with the goal of writing programs which
    can understand human languages.  It is in no way connected to
    Neuro-linguistic programming, which is generally found when searching
    for NLP.
}}

>>>>

\newacronym{RBL}{Real-time Black List}{description={
    Real-time Black Lists are a simple collaborative anti-spam technique
    used to reject or penalise email sent from mail servers reported to
    have sent large volumes of spam.  To use an RBL the mail server makes a
    DNS request incorporating the IP address of the currently connected
    client; if the requested hostname exists the client is on the RBL, and
    the mail server can decide what course of action to take.  XXX
    REFERENCES XXX.
}}

@@

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP is the protocol which transfers mail from the sender to the
    recipient across the Internet.  It is a simple, human readable, plain
    text protocol, making it quite easy to test and debug problems with it.
    A detailed description of \SMTP{} is beyond the scope of this document:
    the original protocol definition is \RFC{}~821~\cite{RFC821}, updated
    in \RFC{}~2821~\cite{RFC2821}; introductory guides can be found
    at~\cite{smtp-intro-01,smtp-intro-02}.
}}

@@

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a powerful method of matching patterns against
    text that are explained in detail in~\cite{Wikipedia-regex, perlre}.
}}

@@

\newacronym{SLCT}{Simple Logfile Clustering Tool}{description={
    SLCT~\cite{slct-paper} is a tool implementing an algorithm designed by
    Risto Vaarandi for identifying and grouping similar log files, then
    producing a regex for each group of lines.
}}

@@

\newacronym{NLP}{Natural Language Processing}{description={
    Natural Language Processing~\cite{Wikipedia-nlp} attempts to increase
    our understanding of the languages normally used by humans (e.g.\
    English, Spanish, Japanese), with the goal of writing programs which
    can understand human languages.
}}

### 2008-05-28 papers/sgai-2008/paper.tex
Lots of corrections after proofreading.

\usepackage[style=altlist,number=none,toc=true]{glossary}
\makeglossary{}

@@

@@

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt existing rules or add new rules to
    parse new inputs, without requiring that user have an in-depth
    knowledge and understanding of the parser's internal workings.  This
    paper presents a novel parsing architecture which aims to make the
    process of parsing new inputs as simple as possible, enabling users to
    trivially add new rules or adapt existing rules (to parse variants of
    existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture provides a
    framework which actions and rules fit into: this framework manages the
    parsing process, provides support functions for the actions, and
    automatically performs some optimisations which improve the speed of
    parsing.
}

@@

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing new measures to supplement the
existing filters where there is a deficiency.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the huge variety of log lines produced by Postfix~\cite{postfix}, the
mail server in use:

@@

    \item The mail administrator can define custom rejection messages,
        resulting in custom log lines.

@@

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.\footnote{A review of the existing parsers
considered for this project will be included in the author's thesis.} The
effort required to adapt and improve an existing parser was judged to be
greater than the effort to write one from scratch, as the techniques used
by the existing parsers severely limited their potential: some ignored the
majority of log lines, parsing specific log lines with great accuracy, but
in an inextensible fashion; others sloppily parsed the majority of log
lines, but were incapable of distinguishing between log lines of the same
type, e.g.\ rejections, and they extracted only a portion of the data
required for this project.

@@

The framework tries each rule in turn until it finds one which matches the
input line, then invokes the action specified by that rule.

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code (significantly lowering the barrier to entry for
new or casual users who need to parse new log lines), and greatly
simplifies the framework, actions and rules. Decoupling the actions from
the framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to deal with the complications
which arise, or the task of reconstructing a mail's journey through
Postfix; actions benefit from having services provided by the framework,
freeing them to concentrate on the task of accurately reconstructing each
mail's journey through Postfix and dealing with the complications arising
(discussed in \sectionref{complications}).

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying and extracting data from log lines;
actions handle the higher level details of following the path a mail takes
through Postfix, assembling the required data, dealing with complications
arising, etc; the framework provides services to actions and stores data.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other.

@@

Actions contain the code required to deal with log lines from a particular
category, e.g.\ rejections.  The actions are parser-specific: each parser
will need to write the actions it requires from scratch, unless it's
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules;
the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{INFO}, the action responsible
for handling informative inputs: these are log lines that supplement the
data gathered from other log lines (such as rejections), but that don't by
themselves represent interesting events.  The third most common action is,
perhaps surprisingly, \texttt{IGNORE}: there are many log lines which are
not of interest to the parser, e.g.\ warning messages about DNS mismatches,
or log lines lacking a way to identify which data structure they should be
associated with.\footnote{Postfix allocates a queueid (explained in the
glossary,~\sectionref{Glossary}) to each mail; most log lines contain the
queueid of the associated mail, but some log lines (particularly with older
versions of Postfix) lack a queueid, preventing their association with a
particular data structure.}  The parser warns about every unparsed log
line, to alert the user that they need to alter or extend their rules; the
\texttt{IGNORE} action does nothing when executed, allowing uninteresting
log lines to be parsed without any effects.  The remaining actions have
only one or two associated rules: in some cases this is because there will
only ever be one type of input for that action, e.g.\ all log lines showing
that a remote client has connected are matched by a single rule and handled
by the \texttt{CONNECT} action; in other cases the action is required to
address a deficiency in the log files, or a complication which arises
during parsing.\footnote{\label{complications footnote}The complications
which arose during development of the parser will be documented in detail
in the author's thesis.}

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to existing actions to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework),
which then check for the extra information and act appropriately.

@@

Rules are responsible for categorising input lines, specifying both the
data to be extracted from each line and the action that should be invoked.
This implementation stores the rules in the same \SQL{} database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [program] The Postfix program whose log lines the rule applies
        to.  This avoids needlessly trying rules which won't match the log
        line, or worse, might match unintentionally.  Rules whose program
        is \texttt{*} will be tried against any log lines which aren't
        parsed by program specific rules, allowing generic rules to be
        written.

    \item [regex] The \regex{} to match the log line against.  For
        efficiency every rule's \regex{} is compiled before attempting to
        parse the log file --- otherwise each \regex{} would be recompiled
        each time it was used, resulting in a large, data dependent
        slowdown.\footnote{This is discussed in more detail in the author's
        thesis.}

    \item [result\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right hand side.

@@

    \item [action] The action that will be invoked when this rule matches a
        log line; actions are described in \sectionref{Actions}.

    \item [hits] This counter is used to optimise the order in which rules
        are tried against the input lines.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more commonly occurring log lines will be tried before
        rules matching less commonly occurring log lines, lowering the
        program's execution time.

@@

\subsubsection{Overlapping rules}

@@

XXX DO I NEED A SECTION ABOUT OVERLAPPING RULES\@?

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse each
rule's regex and determine if any overlap.  The author has not attempted to
do this: given the tremendous power and flexibility of regexes,
particularly Perl's regexes~\cite{perlre}, it would be an extremely
difficult task.} Unintentionally overlapping rules lead to inconsistent
parsing and data extraction because the order in which rules are tried
against each line may change between log files, and the first matching rule
wins.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some log lines and a more general rule to match the
majority, e.g.\ separating \SMTP{} delivery to specific sites from \SMTP{}
delivery to the rest of the world.  The algorithm provides a facility for
ordering overlapping rules: the priority field in each rule (defaults to
zero).  Rules are sorted by priority, highest first, and then rules with
the same priority are sorted by the number of successful matches when
parsing the previous log file.  Negative priorities may be useful for
catchall rules.

@@

    \item [Context-free rules:]  Rules are context free: they do not take
        into account inputs which have previously been seen, nor do they
        perform any kind of lookahead to future inputs.  In context-free
        grammar terms the parser rules could be described as:
        $\text{\textless{}log-line\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

    \item [Partially context-sensitive actions:]  XXX IS THIS THE CORRECT
        TERMINOLOGY\@?  Actions can consult the results of previous actions
        when executing, providing some context sensitivity.  In some cases
        the lack of results from previous actions can be useful
        information, XXX IS THE EXAMPLE NECESSARY\@?  E.g.\ in the Postfix
        parser when a client disconnects if there have been no useful log
        lines for that client the data is discarded, rather than trying to
        store empty records in the database.

    \item [Matching rules to log lines is simple:]  The first matching rule
        wins: there is no backtracking to try alternate rules, no attempt
        is made to pick a \textit{best\/} rule; the first rule which
        matches the line is the rule which parses it and determines the
        action that will be invoked.

    \item [Transduction:]  The parser can be thought of as implementing
        transduction: it takes data in one form (log files) and transforms
        it to another form (a database).  Other implementations are free to
        choose the output format most suitable for their needs.

    \item [Closer to \NLP{} than fixed grammar:]  Unlike traditional
        parsers such as those used when compiling a programming language,
        this architecture does not have a fixed grammar specification which
        inputs must adhere to.  The architecture is designed for easy user
        alteration, in fact such alteration is expected.  XXX NEED MORE
        HERE\@.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the first run discarded, and the
execution time for the remaining 9 runs averaged.  XXX DO I NEED TO EXPLAIN
WHY\@?  The first run is discarded for two reasons:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time and ratio of number of log lines vs execution time (higher
is better).  As the reader can see the ratios are quite tightly banded,
showing that the algorithm scales linearly: the much larger log files
between points 60 and 70 on the X axis in fig.~\refwithpage{execution time
vs file size vs number of lines graph} actually cause the ratio to
increase, rather than decrease.  The strange behaviour where larger log
files are parsed more efficiently is explained fully in \sectionref{Why are
there dips in the graphs?}.

@@

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules (as shown in
fig.~\refwithpage{rule hits graph}).  \showgraph{build/plot-hits}{Hits per
rule}{rule hits graph} Assuming that the distribution of log lines is
reasonably steady over time, program efficiency should benefit from trying
more frequently matching rules before those which match less frequently.
To test this hypothesis three full test runs were performed with different
rule orderings:

@@

    \item [normal]  The most optimal order, according to the hypothesis:
        rules which match most often will be tried first.

@@

\subsubsection{Why are there dips in the graphs?}
\label{Why are there dips in the graphs?}

The dips at log files 22 and 62--68 correspond to peaks in log file size in
fig.~\refwithpage{execution time vs file size vs number of lines graph},
and peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} (where a peak means that more lines are processed per second, i.e.\
performance is better).  The explanation for this took some time to arrive
at, but it turns out to be reasonably simple.  The large log files in
question were caused by a mail forwarding loop, where the distribution of
log lines is quite different to normal, resulting in different performance
characteristics.  Fewer

The vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general \daemon{smtpd} adds the majority of log lines, and also has the
highest number of rules.  \daemon{smtpd} log lines are distributed across
rules much more evenly than the log lines which occur frequently during a
mail loop, so the average number of rules
required to parse a \daemon{smtpd} log line is much higher that the average
number required to parse other log lines.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor}.  When the rule ordering is reversed the majority of log lines
generated by a mail loop will be parsed with very few rules, whereas
without a mail loop the majority of log lines require a large number of
rules; this leads to a noticeable drop in the average time required to
parse a log line, as shown in fig.~\refwithpage{percentage increase of
shuffled and reversed over normal}.  The number of rules which need to be
consulted when the ordering is shuffled varies between the optimum and
nadir, and the performance varies proportionally.

@@

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.\footnote{There is substantially more coverage
discussion in the author's thesis.}  The former is initially more
important, as the parser must successfully parse every line if it is to be
complete, but subsequently the latter takes precedence because reproducing
the path a mail takes through Postfix is the aim of the parser.  Increasing
the percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

@@

>>>>

\usepackage[style=altlist,number=none,toc=true,global=true]{glossary}
\makeglossary{}

@@

% Tell Latex to use scalable fonts
\usepackage{type1cm}

@@

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that user have an in-depth
    knowledge and understanding of the parser's internal workings.  This
    paper presents a novel parsing architecture which aims to make the
    process of parsing new inputs as simple as possible, enabling users to
    trivially add new rules or adapt existing rules (to parse variants of
    existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    as the number of rules and/or size of input increases, making it
    suitable for parsing large corpora or months of accumulated data.
}

XXX DOES THE ABSTRACT NEED TO BE LONGER\@?  IS IT A GOOD TITLE\@?

@@

XXX DO I NEED TO SAY THAT THINGS WILL BE EXPANDED UPON IN MY THESIS\@?

XXX USE citeref AND A REFERENCE PACKAGE WHICH PUTS CLEVER PAGE NUMBERS ON
REFERENCES\@.

XXX CHECK EVERY URL IN THE BIBLIOGRAPHY\@.

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing new measures to supplement the
existing filters where there is a deficiency.  The approach chosen to
measure performance was to analyse the log files produced by the mail
server, rather than modifying the mail server to generate performance
statistics; this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The variety occurs for
several reasons:

@@

    \item The mail administrator can define custom rejection messages,
        resulting in custom log lines~\cite{postfix-lookup-tables,
        smtpd_access_readme, smtpd_per_user_control}.

@@

The parser must be designed so that adding support for parsing new log
lines is a simple task, as the log lines will continue to evolve over time.

Existing parsers were considered, but their parsing coverage was too low,
their parsing too inexact, and/or they did not extract sufficient data for
the purposes of this project.\footnote{A review of the existing parsers
considered for this project will be included in the author's thesis.} The
effort required to adapt and improve an existing parser was judged to be
greater than the effort to write one from scratch, as the techniques used
by the existing parsers severely limited their potential: some ignored the
majority of log lines, parsing specific log lines with great accuracy, but
in an inextensible fashion; others sloppily parsed the majority of log
lines, but were incapable of distinguishing between log lines of the same
type, e.g.\ rejections, and extracted only a portion of the data required
for this project.

@@

The framework tries each rule in turn until it finds one which matches the
input log line, then invokes the action specified by that rule.

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code (significantly lowering the barrier to entry for
new or casual users who need to parse new log lines), and greatly
simplifies the framework, actions and rules. Decoupling the actions from
the framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to perform any tasks specific to
the input being parsed; actions benefit from having services provided by
the framework, freeing them to concentrate on the task of accurately and
correctly processing the information provided by rules.

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying and extracting data from log lines;
actions handle the higher level details of assembling the required data,
dealing with the intricacies of the input being parsed, complications
arising, etc; the framework provides services to actions and manages the
parsing process.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other.  The framework is the core of the
architecture and is deliberately quite simple: the rules deal with the
varying log lines, and the actions deal with the intricacies and
complications encountered when parsing.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches, or log lines lacking a way to identify which data structure
they should be associated with.\footnote{Postfix allocates a queueid
(explained in the glossary,~\sectionref{Glossary}) to each mail; most log
lines contain the queueid of the associated mail, but some log lines
(particularly with older versions of Postfix) lack a queueid, preventing
their association with a particular data structure.}  The parser warns
about every unparsed log line, to alert the user that they need to alter or
extend their rules; the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The remaining actions have only one or two associated rules: in some cases
this is because there will only ever be one type of log line for that
action, e.g.\ all log lines showing that a remote client has connected are
matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during
parsing.\footnote{\label{complications footnote}The complications which
arose during development of the parser will be documented in detail in the
author's thesis.}

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to a single existing
action to work around the difficulties; the remainder were resolved by
adapting actions to pass information to subsequent actions (via the
framework), which then check for the extra information and change their
behaviour appropriately.

@@

Rules are responsible for categorising inputs, specifying both the data to
be extracted from each log line and the action that should be invoked.
This implementation stores the rules in the same \SQL{} database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [program] The Postfix program whose log lines the rule should be
        applied to.  This avoids needlessly trying rules that won't match
        the log line, or worse, might match unintentionally.  Rules whose
        program is \texttt{*} will be tried against any log lines not
        parsed by program specific rules, allowing generic rules to be
        written.

    \item [regex] The \regex{} to match the log line against.  For
        efficiency every rule's \regex{} is compiled before attempting to
        parse the log file --- otherwise each \regex{} would be recompiled
        each time it was used, resulting in a large, data dependent
        slowdown.\footnote{This will be discussed in more detail in the
        author's thesis.}  XXX MAYBE SCRAP THE SECOND SENTENCE\@?

    \item [result\_cols] Specifies how the fields in the
        log line will be extracted.  The format is:
        \tab{} \texttt{smtp\_code = 1; recipient = 2, sender = 4;}
        \newline i.e.\ semi-colon or comma separated assignment statements,
        with the variable name on the left and the matching capture from
        the \regex{} on the right.

@@

    \item [action] The action that will be invoked when this rule matches a
        log line.

    \item [hits] This counter is used to optimise the order in which rules
        are tried against the input lines.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more frequently will be tried before rules matching less
        frequently, lowering the parser's execution time.

@@

@@

XXX DO I NEED TO DISCUSS OVERLAPPING RULES\@?

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse each
rule's regex and determine if any overlap.  The author has not attempted to
do this: given the tremendous power and flexibility of regexes,
particularly Perl's regexes~\cite{perlre}, it would be an extremely
difficult task.} Unintentionally overlapping rules lead to inconsistent
parsing and data extraction because the first matching rule wins and the
order in which rules are tried against each line may change between log
files.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some log lines and a more general rule to match the
majority, e.g.\ separating \SMTP{} delivery to specific sites from \SMTP{}
delivery to the rest of the world.  The algorithm provides a facility for
specifying the order of overlapping rules: the priority field in each rule
(defaults to zero).  Rules are sorted by priority, highest first, and then
rules with the same priority are sorted by the number of successful matches
when parsing the previous log file.  Negative priorities may be useful for
catchall rules.

@@

    \item [Context-free rules:]  Rules are context free: they do not take
        into account inputs which have previously been seen, nor do they
        perform any kind of lookahead to future inputs.  In context-free
        grammar terms the parser rules could be described as:
        $\text{\textless{}input\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

    \item [Partially context-sensitive actions:]  XXX IS THIS THE CORRECT
        TERMINOLOGY\@?  Actions can consult the results (or lack of
        results) of previous actions when executing, providing some context
        sensitivity.  Actions can not inspect past or upcoming inputs.

    \item [Matching rules to log lines is simple:]  The first matching rule
        wins: there is no backtracking to try alternate rules, no attempt
        is made to pick a \textit{best\/} rule; the first rule which
        matches the line is the rule which parses it and determines the
        action that will be invoked.  The priority attribute may be used to
        change the ordering of rules.

    \item [Transduction:]  The parser can be thought of as implementing
        transduction: it takes data in one form (log files) and transforms
        it to another form (a database); other output formats may be more
        suitable for other implementations.

    \item [Closer to \NLP{} than fixed grammar:]  Unlike traditional
        parsers such as those used when compiling a programming language,
        this architecture does not have a fixed grammar specification which
        inputs must adhere to.  The architecture is capable of dealing with
        overlapping inputs, out of order inputs, and ambiguous inputs where
        heuristics must be applied --- all have arisen and been
        successfully dealt with in the Postfix log parser.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the results from first run discarded,
and the execution time for the remaining 9 runs averaged.  XXX DO I NEED TO
EXPLAIN WHY\@?  The first run is discarded for two reasons:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB, and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time, and ratio of number of log lines vs execution time (higher
is better in both cases).  As the reader can see the ratios are quite
tightly banded across the graph, showing that the algorithm scales
linearly: the much larger log files between points 60 and 70 on the X axis
in fig.~\refwithpage{execution time vs file size vs number of lines graph}
actually cause the ratio to increase (i.e.\ improve), rather than decrease.
The strange behaviour where larger log files are parsed more efficiently is
explained in \sectionref{Why are there dips in the graphs?}.

@@

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules in a Power Law
distribution~\cite{powerlaw} (as shown in fig.~\refwithpage{rule hits
graph}).  \showgraph{build/plot-hits}{Hits per rule}{rule hits graph}
Assuming that the distribution of log lines is reasonably steady over time,
parser efficiency should benefit from trying more frequently matching rules
before those which match less frequently.  To test this hypothesis three
full test runs were performed with different rule orderings:

@@

    \item [normal]  Hypothetically the most optimal order: rules which
        match most often will be tried first.

@@

\noindent\textbf{Why are there dips in the graphs?}

\label{Why are there dips in the graphs?}

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{execution time vs file size vs number of
lines graph}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that more lines are processed per
second, i.e.\ performance is better).  The explanation for this took some
time to arrive at, but it turns out to be reasonably simple.  The large log
files in question were caused by a mail forwarding loop, where the
distribution of log lines is quite different to normal, resulting in
different performance characteristics.\footnote{The cause of the mail
forwarding loop will be explained in the author's thesis.}

XXX REWRITE THIS NEXT PARAGRAPH SO IT DOESN'T MENTION postfix/smtpd.  The
vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general \daemon{smtpd} adds the majority of log lines, and also has the
highest number of rules.  \daemon{smtpd} log lines are distributed across
rules much more evenly than the log lines which occur frequently during a
mail loop, so the average number of rules required to parse a
\daemon{smtpd} log line is much higher that the average number required to
parse other log lines.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{percentage increase of shuffled
and reversed over normal}.  When the rule ordering is reversed the majority
of log lines generated by a mail loop will be parsed using very few rules,
because there are very few rules for the Postfix components producing the
log lines.  Without a mail loop the Postfix components producing the
majority of log lines have a large number of rules associated with them,
and reverse ordering deliberately starts with the least likely rules and
moves towards the most likely.  This leads to a noticeable drop in the
average time required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{percentage increase of shuffled and reversed over
normal}.  The number of rules which need to be consulted when the ordering
is shuffled varies between the optimum and the nadir, and the performance
varies proportionally.

@@

XXX THIS NEEDS TO BE IMPROVED AND PROOFREAD\@.

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.\footnote{There will be substantially more
coverage discussion in the author's thesis.}  The former is initially more
important, as the parser must successfully parse every line if it is to be
complete, but subsequently the latter takes precedence because reproducing
the path a mail takes through Postfix is the aim of the parser.  Increasing
the percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

@@

XXX WRITE THE CONCLUSION\@.
